---
title: "wimbo wave music recommender"
author: "karani"
format: html
editor: visual
---

Import the packages to use

```{r}
library(tidyverse)
```

import the data set from tidy Tuesday github page

```{r}
polls <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/polls.csv')

rankings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/rankings.csv')

# view the data
View(polls)
```

```{r}
#View the data
View(rankings)
```

Basic EDA

```{r}
# to view waht song was on the participiating critics top lists
polls %>% 
  count(title, sort = TRUE)
```

curious to know was there a golden age of rap according to our data?

```{r}
polls %>% 
  count(year) %>% 
  mutate(decade = floor(year / 10) * 10) %>%
  mutate(decade = as.factor(decade)) %>% 
  ggplot(aes(x = year, y = n, fill = decade)) + geom_col()
```

aggregated decade summary by decade

```{r}
polls %>% 
  count(year) %>% 
  mutate(decade = floor(year / 10) * 10) %>%
  mutate(decade = as.factor(decade)) %>% 
  ggplot(aes(x = decade, y = n, fill = decade)) + geom_col()

# we observe that the 90s are the golden days of rap
```

let us count who is the most famous artist from the polls data

```{r}
polls %>% 
  count(artist, sort = TRUE) %>% 
  ggplot(aes(x = n)) + geom_density()

```

sum of the songs artist appear: remove featuring aspect of an artist in a song

```{r}
rankings %>% 
  select(artist, n, n1, n2, n3, n4, n5) %>% 
  group_by(artist) %>% 
  summarise_all(sum) %>% 
  filter(!str_detect(artist, "ft.")) %>% 
  ggplot(aes(x = n1, y = n5 )) +
  geom_jitter()
```

artists with the top rank 1 songs (n1)

```{r}
rankings %>% 
  select(artist, n, n1, n2, n3, n4, n5) %>% 
  group_by(artist) %>% 
  summarise_all(sum) %>% 
  filter(!str_detect(artist, "ft.")) %>% 
  arrange(desc(n1)) %>% 
  slice(1:5)
```

```{r}
rankings %>% 
  select(artist, n, n1, n2, n3, n4, n5) %>% 
  group_by(artist) %>% 
  summarise_all(sum) %>% 
  filter(!str_detect(artist, "ft.")) %>% 
  arrange(desc(n5)) %>% 
  slice(1:5) 
```

15% of the top songs were voted by a single country. does that allude to a country bias by the critics?

```{r}
polls %>% 
  count(title, critic_country, name = "song_nom") %>% 
  add_count(title, name = "number_of_countries") %>% 
  filter(number_of_countries == 1, critic_country != "us") %>% 
  nrow() / nrow(polls)
  
  
```
